---
title: "Macroinvertebrate Data Analysis"
author: "Your Namge"
date: "October  , 2022"
format: pdf
---

## Getting Setup

Before getting started, it is a good practice to start with a clean slate by restarting your R session

Go to the top menu, select Session then select Restart R

Keep all of your analysis in this same .qmd file and keep all the associated files in the project folder. Make sure the project name is showing in the upper right of the RStudio window.

# Load packages

```{r setup}

library(tidyverse) ## for readr, dplyr, ggplot2
library(vegan) ## for calculating diversity measures (you can remove this if you are not calculating diversity measures)
library(ggbeeswarm) ## for jittering points
library(flextable) ## for making formatted tables that can be pasted into Word or Google Docs

```

# Import Data

Figure out which data file(s) you need to import. Find the code you need in the "importing_data" folder within the "script_templates" folder. Copy and paste the code you need in the chunk below.

```{r import data}


```

# Data Wrangling

Find the code you need to prepare or "wrangle" your data to make any calculations needed and combine any datasets you need to combine by looking in the "data_wrangling" folder within the "script_templates" folder. Copy and paste the code you need in the chunk below.

```{r data wrangling}



```

# Data Description

Make a formatted table of the data that you used in your analysis to include in the Data Analysis Methods section of your paper. Make sure to remove variables you don't need (if you are only looking at one season, you can remove season from this table)

```{r table of data used}

#To get a formatted table, make sure that the {flextable} package is loaded
#If you want to format the table yourself, then erase the code from 
#where it says "#create the formatted table" on

table.sum <- ___ %>%  #put the name of your wrangled data here
  
  #summarize by date, year, season, location (to get number of microhabs sampled per date)
  group_by(date, year, season, location) %>% 
    summarize(
      area = sum(benthicArea),
      mon.ADD = mean(mon.ADD, na.rm=TRUE),
      mon.precip = mean(mon.precip, na.rm=TRUE),
      samples = n()
    ) %>% 
  
  #change date to month/day
  mutate(date = format(as.Date(date), '%m/%d')) %>% 


  #summarize by year, season, location 
  group_by(year, season, location) %>% 
    summarize(
      dates = str_c(date, collapse = ", "),
      samples = sum(samples),
      area = sum(area),
      mon.ADD = round(mean(mon.ADD, na.rm=TRUE), digits=0),
      mon.precip = round(mean(mon.precip, na.rm=TRUE), digits=0)
    )  
  

#create the formatted table
ft <- flextable(table.sum)

#create header labels
ft <- set_header_labels(ft,
                        year = "Year",
                        season = "Season",
                        location = "Location",
                        dates = "Dates",
                        samples = "# of Samples",
                        area = "Total Area Sampled (m^2)",
                        mon.ADD = "Monthly ADD",
                        mon.precip = "Monthly Precipitation (mm)"
                        )


#bold the headings
ft <- bold(ft, part="header")

ft <- merge_v(ft, j = ~ year + season)

#set table properties
ft <- set_table_properties(ft, layout = "autofit", width = 1)

#center columns
ft <- align(ft, align = "center", part = "all" )


#add lines between year x season
ft <- theme_vanilla(ft) %>% 
  fontsize(size = 10)

ft <- fix_border_issues(ft)

#print table
ft

#send to MS Word if desired
#comment out or delete the line below before you render a quarto document
#print(ft, preview = "docx") # can change this to "html" or "pptx"

#see https://davidgohel.github.io/flextable/ for more information and formatting options

```

Use the template code below to calculate the descriptive statistics for your variable of interest. The output of this code should not all be reported in your paper. Decide which (if any) of the descriptive statistics you should report and make sure to round the values. Descriptive statistics can be reported in the text of the results (see this [example results section](https://docs.google.com/document/d/1vwyOSmhZzTuIfAEbpeiv3yCsU1UMZDsTXgnbE80nPTU/edit?usp=sharing)) if relatively simple, or you can make a formatted table of just the necessary, rounded values to include in your paper (in which case you might want to use "formatted_descriptiveStats_table.R" in the "data_description" script templates to help you make a formatted table).

```{r descriptive stats}

# the code below will calculate descriptive statistics for a variable of interest grouped by another variable

df.sum <- ___ %>% # put the name of the data frame here
  filter(!is.na(___)) %>% # remove missing values from the variable of interest
  group_by(___) %>% # put the name of the grouping variable(s) here
  summarize(mean = mean(___), # put the name of the variable you want to summarize in this & following blanks
            median = median(___), 
            SD = sd(___), 
            IQR = IQR(___), 
            min = min(___),
            max = max(___),
            n())

df.sum

```

Make a formatted table of the data that you used in your analysis to include in the Data Analysis Methods section of your paper.

```{r table of data used}



```

# Data Visualization

Find the code you need to visualize (=graph) your data by looking in the "graphing" folder within the "script_templates" folder. Copy and paste the code you need in the chunk below.

```{r graphs}



```

To save your graph with the `ggsave()` function, you need to name the resulting file with surrounding " ", and indicate the size via height, width, and units. Don't forget to save the graph with a dpi call between 300-500 to make it nice and crisp! Look at the `ggsave()` help file for more information and options.

```{r Save your graph, eval = F}

# save the graph!
ggsave(filename="___.png",  #recommended to use.png or .jpg file types
       height = 7, width = 8, units = "in", 
       dpi = 300)

```

# Citing R in Your Paper

In the Data Analysis Methods section of your paper, you need to cite how the data were analyzed including (1) any calculations done (for example, how was density calculated?), and (2) what software was used for the analysis.

Here is a detailed explanation of How to Cite R in your Methods.

You need to cite base R but how will you know what version of R you are using?

You should also cite the most important packages used.

All of you should cite the {tidyverse} super package since we used {reader} and {dplyr} to import and wrangle our data. If you used the dot plot in your paper, then you used {ggplot2} which is part of the {tidyverse} (so you don't need to cite it separately), but also {ggbeeswarm} which is NOT part of the {tidyverse} so it needs to be cited separately. If you used the formatted table in your paper, then cite the {flextable} package.

To properly cite all the packages listed above, you need to find out the information needed for each package. Run the code below to determine the version of R and packages used.

```{r sessionInfo}

sessionInfo()

```

The above is the information you need to cite which version of R as well as any packages you used. While the version number is necessary, you also need more information to cite R and the packages. Refer to [How to Cite R in your Methods](https://docs.google.com/presentation/d/1RG4_R-MDOy1vbMz-M1gfOcMtgqWu9Kvv_R4ex5QdcVk/edit?usp=sharing) to help you figure out how to cite in the text as well as the full citation in the Literature Cited section at the end of your paper.

Use the code below to get more necessary information to cite R itself

```{r cite R}

#citation info for base R 
citation()

```

Use the code below to cite the {tidyverse} package and then repeat the code replacing `tidyverse` with any other packages you need to cite

```{r cite tidyverse}

#citation info for base R 
citation("tidyverse")

```

## Rendering

Don't forget to "render" your .qmd file when you are done. Make sure that each code chunk has an unique name or no name before you render. Make sure all your files are in the project folder.

Render straight to PDF or render to HTML, open the HTML file in your browser, and then "print" to a PDF file.

Insert a file printout of the PDF in the "Project Data Analysis" section of your Lab Notebook.
